# Data Documentation

The dataset can be found here: [Kaggle College Basketball Dataset](https://www.kaggle.com/datasets/andrewsundberg/college-basketball-dataset).

[**First 10 Rows of CBB Dataset:**]{.underline}

```{r}
#| echo: FALSE
#| warning: FALSE
#| label: tbl-cbb
#| tbl-cap: "College Basketball Defensive Data for NCAA Basketball Seasons from 2013-2025"
#| results: asis

cbb <- read.csv("cbb.csv")

writeLines("\\footnotesize")
knitr::kable(head(cbb[c(1:4,6:7,9,22:24)], 10))
writeLines("\\normalsize")

```

[**Variable List:**]{.underline}

```{r}
#| echo: FALSE
#| warning: FALSE
#| label: tbl-var
#| tbl-cap: "Dataset Variable Names and Descriptions"
#| results: asis

cbb.variables <- data.frame(Variable=c("TEAM", "CONF", "W", "ADJDE", "BARTHAG", 
                                       "POSTSEASON", "SEED", "YEAR"),
                        Description=c("NCAA Division I School Name",
                                      "Athletic Conference", 
                                      "Games Won", 
                                      "Adjusted Defensive Efficiency (Pts Allowed/100 Poss.)",
                                      "Power Rating (Win % vs avg DI team)",
                                      "Tournament Exit Round (e.g., S16, F4, Champion)",
                                      "Tournament Seed", "Season Year"))

writeLines("\\footnotesize")
knitr::kable(cbb.variables)
writeLines("\\normalsize")

```

------------------------------------------------------------------------

**Population of Interest**: All Division I College Basketball Schools that played Division I Basketball between 2013-2025. As seen in @tbl-cbb

**Variable of Interest**:

*cbb\$ADJD* → Adjusted Defensive Rating. According to the original data source, this variable is "\[an\] estimate of the defensive efficiency (points allowed per 100 possessions) a team would have against the average Division I offense" (Sundberg, 2026).

------------------------------------------------------------------------

**Research Question**: What is the maximum defensive rating needed to be a true national championship contender in March Madness?

**Reasoning**: The maximum value would indicate the level of defense needed for a college basketball team to have realistic expectations of winning the entire tournament, because the higher the defensive rating, the worse a team is defensively. It would be helpful for coaches and players to know what this threshold is to manage expectations and to provide a framework of where they need to get to defensively in order to advance further in the tournament.

**Hypothesized Distribution**: We believe that this scenario would fit an exponential distribution, because defensive rating is a non-negative, continuous variable. However, it would be a shifted exponential because an extremely low season-long defensive rating is near impossible for teams today. We also believe that a maximum is appropriate because most champions are going to have a lower defensive rating, as it shows that the team is less likely to give up points on opposing possessions. A maximum value in this case would represent the highest defensive rating a team can have before they are no longer considered a national championship contender.

**CDF and pdf**:

\$\$

F_X(x\|\beta) =

\begin{cases} 
      1-e^{-\frac{x-L}{\beta}} & x \geq L \\
      0 & x<L\\
\end{cases}

$$

$$

max(X_1,...,X_n)=T(X_1,...,X_n)\stackrel{ind}{=}P(X_1\le t)P(X_2\le t)...P(X_n\le t)

\$\$

[pdf:]{.underline}

\$\$

f_X(x\|\beta)=\frac{1}{\beta}e\^{-\frac{x-L}{\beta}}, x\>0

\$\$
